Estimating Neighborhood Rents with Scraped Data

### 2018-06-05 manuscript/index.Rmd
progress through 6/4

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess"
date: "June 2, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: "header.html"
      after_body: "../footer.html"
abstract: "Online classified ads are a key source of information about local rental housing markets. This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources. The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models. All models' posterior distribution are estimated with integrated nested Laplace approximations (INLA). Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 3rd. The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $95 for the training data and $140 for the test data."
---

@@

```{r setup, include=FALSE}
#packages
library(tidyverse) #need dev version of ggplot
library(forcats)
library(haven)
library(lubridate)
library(haven)
library(zoo)
library(ggthemes)
library(RColorBrewer)
library(sp)
library(spdplyr)
library(geosphere)
library(rgdal)
library(rgeos)
library(sqldf)
library(spdep)
library(viridis)
library(latex2exp)

@@

[American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more. This publication lag for ACS data prohibits analyzing any market dynamics beyond long-term trends, particularly for estimates of smaller areas of aggregation like neighborhoods where ACS uses 5-year rolling-average methods as well. While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties. This would-be problem for academic researchers creates an opportunity for applying web scraping methods to generate a novel data on housing availability and costs.

This study demonstrates why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography. I draw on a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA for this research objective. Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data. The smoothing models use the panel of quarterly neighborhood estimates' spatial and temporal structure to understand the underlying generative process, finding that a combination of local smoothing and autoregression (order 1) predict future levels with an average error of about $140 or 8%. In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.

@@

Existing public data are essentially limited to the American Community Survey (ACS) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html) following the cessation of long-form decennial censuses in 2000. These data theoretically have two sources of error relative to the ``true'' current median rent for a market and  neighborhood approximations like census tracts. Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication (i.e. ACS estimates are usually at least 2 years old, decennial many more). This feature of Census data is salient given newer housing units' generally higher rents and greater-than-average vacancy, both factors which lead to underestimating the present levels of rents for available housing units when either the ACS or decennial census are used as a current indicator. For already-leased rental housing units that are sampled, the temporal lag generates measurement error such that data on a given building's current asking rents may be underestimated relative to current tenants' payment levels.

@@

![**Figure 1.** Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)

@@

Figure 2 shows a heatmap for the time-series of one-bedroom listings per day. The grey cells indicate days where our scraper was offline. Outages were more common in the earliest phases of work, but now are relatively rare.

At the end of February 2018 we implemented a procedure to have the Helena program immediately restart scraping (i.e. re-query the available listings as a new index to work through) in order to ensure any and all listings are scraped each day. The net impact of this change is fairly evident in the time-series heatmap. We now see more temporal structure to the listing count for each day---relatively fewer listings are posted on Sundays, though this could be related to greater listing on Saturdays (i.e. everything is posted at the beginning of the weekend in order to maximize weekend housing searchers' likelihoods of seeing it). Perhaps unsurprisingly, there were dips in listing activity on prominent holiday's like New Year's Day, Easter and Memorial Day.

@@

![**Figure 2.** Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series.](../output/n_heatmap.png)

@@

![**Figure 3.** Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018. Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual (March and September), and ACS are annual (i.e. respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)

Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market. Both data suggest continued increases in the marketwide median rent into Fall 2017, though the Craigslist slope is about twice in its size. Some of the spike observed between the second and third data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate upward in favor of regularly posted commerical listings. Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability. This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.

@@

The focus of the data description now turns towards estimating median rents for specific neighborhoods. While zip codes, unofficial City neighborhoods and even arbitrary grid geographies can all be used to aggregate (or in a graphical sense, rasterize) micro-data about rental housing,  . Figure 4

[Reference 2000 Census tract map for Seattle](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf)

[Reference 2010 census tract map for Seattle](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf)

@@

![**Figure 4.** Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](../output/rent_map_across_1B.png)

@@

@@

![**Figure 5.** Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](../output/rent_scatter_across_1B.png)

@@

@@

- Need to build tables from output
- Need to do write up for tables
- Need to add graphic for predicted values

@@

summary(m.int)
sea_df$int_Med <- m.int$summary.linear.predictor[, "0.5quant"]
sea_df$int_SD <- m.int$summary.linear.predictor[, "sd"]
sea_df$int_postWidth <- m.int$summary.linear.predictor[, "0.975quant"] - m.int$summary.linear.predictor[,"0.025quant"]

@@

summary(m.ns)
sea_df$ns_Med <- m.ns$summary.linear.predictor[, "0.5quant"]
sea_df$ns_SD <- m.ns$summary.linear.predictor[, "sd"]
sea_df$ns_postWidth <- m.ns$summary.linear.predictor[, "0.975quant"] - m.ns$summary.linear.predictor[,"0.025quant"]
sea_df$ns_Eff <- m.ns$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
sea_df$nsar1_Med <- m.nsar1$summary.linear.predictor[, "0.5quant"]
sea_df$nsar1_SD <- m.nsar1$summary.linear.predictor[, "sd"]
sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
sea_df$nsar1_Eff <- m.nsar1$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

summary(m.bym)
sea_df$bym_Med <- m.bym$summary.linear.predictor[, "0.5quant"]
sea_df$bym_SD <- m.bym$summary.linear.predictor[, "sd"]
sea_df$bym_postWidth <- m.bym$summary.linear.predictor[, "0.975quant"] - m.bym$summary.linear.predictor[,"0.025quant"]
sea_df$bym_Eff <- m.bym$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

summary(m.spt)
sea_df$spt_Med <- m.spt$summary.linear.predictor[, "0.5quant"]
sea_df$spt_SD <- m.spt$summary.linear.predictor[, "sd"]
sea_df$spt_postWidth <- m.spt$summary.linear.predictor[, "0.975quant"] - m.spt$summary.linear.predictor[,"0.025quant"]
sea_df$spt_Eff <- m.spt$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

>>>>

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess"
date: "June 2, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: "header.html"
      after_body: "../footer.html"
abstract: "Online classified ads are a key source of information about local rental housing markets. This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources. The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models. All models' posterior distribution are estimated with integrated nested Laplace approximations (INLA). Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 3rd. The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $100 for the training data and $140 for the test data."
---

@@

```{r setup, include=FALSE}
#packages
library(tidyverse) #need dev version of ggplot
library(forcats)
library(haven)
library(lubridate)
library(haven)
library(zoo)
library(ggthemes)
library(RColorBrewer)
library(sp)
library(spdplyr)
library(geosphere)
library(rgdal)
library(rgeos)
library(sqldf)
library(spdep)
library(viridis)
library(latex2exp)
library(plotly)

@@

Census data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more. This publication lag for these data prohibits analyzing market dynamics beyond long-term trends, particularly for smaller areas of aggregation like neighborhoods where the estimate may be an average of survey responses from a 5 year period. While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties. This would-be problem for academic researchers creates an opportunity for applying web scraping methods to generate a novel data on housing availability and costs.

This study demonstrates why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography. I draw on a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA for this research objective. Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data. The models use a quarterly panel of observed neighborhood median one-bedroom rents and produce temporally and spatially smoothed predictions about neighborhood rent levels. The best model specification in terms of forecasting Q2 2018 based on Q1 2017 and Q1 2018 has a combination of local spatial smoothing and autoregression (order 1), with an average forecast error of about $140 or 8% for what. In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.

@@

Existing public data are essentially limited to the [American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html) following the cessation of long-form decennial censuses in 2000. These data theoretically have two sources of error relative to the ``true'' current median rent for a market and  neighborhood approximations like census tracts. Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication (i.e. ACS estimates are usually at least 2 years old, decennial many more). This feature of Census data is salient given newer housing units' generally higher rents and greater-than-average vacancy, both factors which lead to underestimating the present levels of rents for available housing units when either the ACS or decennial census are used as a current indicator. For already-leased rental housing units that are sampled, the temporal lag generates measurement error such that data on a given building's current asking rents may be underestimated relative to current tenants' payment levels.

@@

![**Figure 1.** _Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right)._](../output/map_point_density.png)

@@

Figure 2 is a heatmap for the time-series of one-bedroom listings per day. The grey cells indicate days where our scraper was offline. Outages were more common in the earliest phases of work, but now are relatively rare.

At the end of February 2018 we implemented a procedure to have the Helena program immediately restart scraping (i.e. re-query the available listings as a new index to work through) in order to ensure any and all listings are scraped each day. The net impact of this change is fairly evident in the time-series heatmap. We now see more temporal structure to the listing count for each day---relatively fewer listings are posted on Sundays, though this could be related to greater listing on Saturdays (i.e. everything is posted at the beginning of the weekend in order to maximize weekend housing searchers' likelihoods of seeing it). Perhaps unsurprisingly, there were dips in listing activity on prominent holidays like Christmas, New Year's Day, and Memorial Day.

@@

![**Figure 2.** _Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series._](../output/n_heatmap.png)

@@

![**Figure 3.** _Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018. Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual (March and September), and ACS are annual (i.e. respective 1-Year Estimates for each year period)._](../output/rent_lineplot.png)

Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market. Both data suggest continued increases in the marketwide median rent for the first half of 2017, though the Craigslist slope is about twice in its size. Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commerical listings. Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability. This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.

@@

The focus of the data description now turns towards estimating median rents for specific neighborhoods. While zip codes, unofficial City neighborhoods and even arbitrary grid geographies can all be used to aggregate micro-data about rental housing, we compare census tract estimates because they roughly approximate neighborhood-size units and can be linked with a variety of other data in Seattle and more generally. The following links provide reference maps for Seattle Census tracts ([2000](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf) and [2010](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf))

In the following graphic, the left most pane shows median one-bedroom asking rent estimates using data collected from Craigslist in 2017-2018. The middle pane shows Dupre+Scott tract median 1B contract rent estimates for one-bedrooms in 20+ unit apartment structures according to their September 2017 (Fall) survey of local landlords and property managers. These data were only available in 2000 Census geography, so the neighborhood units differ slightly (124 tracts versus 134 tracts) in this map compared to the other two. Though less than ideal, this limitation does not drastically alter the comparison between data.  Finally, on the right are ACS 2012-2016 5-Year Estimates (the most recent ACS small-area data product available). In all map panes, grey cells denote missing values.

@@

![**Figure 4.** _Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled._](../output/rent_map_across_1B.png)

Both the Craigslist and Dupre+Scott data largely agree on the spatial distribution of one-bedroom rents. The similarity is greatest in the downtown and higher-density neighborhoods where the stock of corporate-held rental housing units theoretically accounts for a most of a tract's rental housing units (i.e. Dupre+Scott closer to a census if they sample all 20+ unit properties). Nonetheless, the systematic difference is such that Craigslist levels are typically at least a little higher. A detailed comparison of the Craigslist and Dupre+Scott estimates follows in the next figure where we directly compare 2000-definition census tract estimates from Craigslist and Dupre+Scott.

Though the spatial distribution of the ACS estimates resembles that of the other two data sources, the values are nonethless systematically lower (in most cases by hundreds of dollars). The ACS measures therefore give insight into the relative ranking of neighborhoods in terms of rental costs, but are limited for studying neighborhood dynamics outside of long-run trend.

@@

Figure 5 shows a scatterplot/bubbpleplot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist. The purple dashed line indicates the loess smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and good agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

@@

![**Figure 5.** _Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties. Purple dashed line indicates loess fit._](../output/rent_scatter_across_1B.png)

An alternative explanation fpr the difference between measures comes in considering the units these properties are sampling from in a given neighborhood. The greater difference between the measures among lower-rent neighborhoods (with Craigslist estimating these as more expensive than Dupre+Scott) provides some evidence for this theory. These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties. In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf).

To the extent that these smaller property units may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone. Similarly, the more that smaller properties account for the overall population of rental housing units, the more likely that higher values indicated in Craigslist are not erroneous but instead reflective of a broader set of residential units.

@@

The following models learn from the Craigslist data accumulated up till the prior quarter (i.e. `r as.yearqtr(Sys.Date())-1`) to accomplish a few different goals. The most prominent goal is to use the panel data's temporal and spatial structure to understand variation in rent levels over time and across space. A secondary goal is to forecast the Seattle neighborhood rent surface for the current quarter. Finally, the following models allow for relatively straightforward imputation of missing values based on time period and the rent values of proximate tracts (though there are only two tracts where we have never observed a one-bedroom listing).

I assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and estimate the posterior distribution of parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) with the default, weakly-informative priors for all random-effects.

Formally, models assume:

$log(Y_{it})|\theta_{it} \sim \mbox{Normal}(\theta_{it})$ with:

Model                    | Abbr.   | Formula
------------------------ | ------- | ----------------------------------------------
Fixed Intercept          | int     | $\theta_{it} = \beta_0 + Qtr$
Non-Spatial              | ns      | $\theta_{it} = \beta_0 + \epsilon_i + Qtr$
Non-Spatial AR(1)        | nsar1   | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i$
BYM Spatial AR(1)        | bym     | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i.$
BYM Spatiotemporal AR(1) | spt     | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i. + \epsilon_{it}$

The Fixed Intercept model (int) uses fixed effects for quarters only. The Non-Spatial (ns) model adds random noise (i.i.d.) tract random-effects. The Non-Spatial AR(1) model (nsar1) uses AR(1) process to model the temporal dynamics (formally, $\epsilon_1 \sim \mbox{Normal}(0, (\tau(1-\rho^2))^{-1}))$ where $\epsilon_t = \rho x_{t-1} + \epsilon_t$). This model includes a random noise term for each quarter too.

The BYM Spatial with AR(1) is the [Besags-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466) intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure. The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $S_i|S_j, j \in ne(i) ~ \mbox{Normal}(\bar{S_i}, \frac{\sigma^{2}_s}{m_j})$. Lastly, the BYM Spatiotemporal AR(1) (spt) adds an exchangeable random-effect for each period to allow for area-specific deviations in time trends.

The space and time smoothing models used in this section partially pool the data (e.g. the same observation at different points in time, a tract's neighboring units) to measure the degree of variation across different quarters and at different points in space. These models are common in epidemiology and disease-mapping applications where researchers need to estimate a quantity of interest like a relative risk but have sparse data for some areas (which may differ from truth due to sampling variability). Modelling a panel's error structure with random-effects for space and time penalizes bumpiness (local deviations) in predicted values and theoretically improves out-of-sample accuracy (because large local deviations can stem from low $n$ in a given area). The spatial models in particular allow for missing data to be imputed based off of a local average of neighbors rather than the global or period-specific average.

@@

Visit the interactive results dashboard [here](http://hesscl.com/smooth-sea/dashboard/index.html).

@@

#summary(m.int)
sea_df$int_Med <- m.int$summary.linear.predictor[, "0.5quant"]
sea_df$int_SD <- m.int$summary.linear.predictor[, "sd"]
sea_df$int_postWidth <- m.int$summary.linear.predictor[, "0.975quant"] - m.int$summary.linear.predictor[,"0.025quant"]

@@

#summary(m.ns)
sea_df$ns_Med <- m.ns$summary.linear.predictor[, "0.5quant"]
sea_df$ns_SD <- m.ns$summary.linear.predictor[, "sd"]
sea_df$ns_postWidth <- m.ns$summary.linear.predictor[, "0.975quant"] - m.ns$summary.linear.predictor[,"0.025quant"]
sea_df$ns_Eff <- m.ns$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

#summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
sea_df$nsar1_Med <- m.nsar1$summary.linear.predictor[, "0.5quant"]
sea_df$nsar1_SD <- m.nsar1$summary.linear.predictor[, "sd"]
sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
sea_df$nsar1_Eff <- m.nsar1$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

#summary(m.bym)
sea_df$bym_Med <- m.bym$summary.linear.predictor[, "0.5quant"]
sea_df$bym_SD <- m.bym$summary.linear.predictor[, "sd"]
sea_df$bym_postWidth <- m.bym$summary.linear.predictor[, "0.975quant"] - m.bym$summary.linear.predictor[,"0.025quant"]
sea_df$bym_Eff <- m.bym$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

#summary(m.spt)
sea_df$spt_Med <- m.spt$summary.linear.predictor[, "0.5quant"]
sea_df$spt_SD <- m.spt$summary.linear.predictor[, "sd"]
sea_df$spt_postWidth <- m.spt$summary.linear.predictor[, "0.975quant"] - m.spt$summary.linear.predictor[,"0.025quant"]
sea_df$spt_Eff <- m.spt$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

```{r bym_lineplot, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
bym_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = bym_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "E. Spatial AR(1)") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

ggplot(bym_df %>% filter(model == "E. Spatial AR(1)"), 
         aes(x = listingQtr, y = value, group = GISJOIN, color = qtrDev, label = GISJOIN)) +
  geom_line(lwd = .25, color = "black", alpha = .5) +
  geom_point() +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_viridis_c(option = "A", labels = scales::dollar) +
  xlab("") +
  ylab("Median 1B Rent\n") +
  labs(color="Quarterly\nDeviation") +
  theme_minimal() +
  ggsave(filename = "../output/bym_lineplot.png",
         width = 6, height = 4, dpi = 500)
```

![**Figure 6.** _Lineplot of Spatial AR(1) model predicted values._](../output/bym_lineplot.png)

```{r avp_map_series, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
obs_map_series <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent)) +
  facet_grid(~ listingQtr) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80",
                       limits = c(800, 2800), option = "A") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        legend.position = "top") +
  guides(fill = "none") +
  labs(subtitle = "Observed", fill = "1B Rent")

bym_map_series <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = bym_Med)) +
  facet_grid(~ listingQtr) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80",
                       limits = c(800, 2800), option = "A") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        legend.position = "bottom",
        legend.key.height = unit(.1, "inch")) +
  guides(fill = guide_colourbar(barwidth = 15,
       label.position = "bottom")) +
  labs(subtitle = "Spatial AR(1) Prediction", fill = "1B Rent")

obs_map_series + bym_map_series +
  plot_layout(ncol = 1) +
  ggsave(filename = "../output/avp_map_series.png",
         width = 6, height = 5, dpi = 500)
```

![**Figure 7.** _Chloropleth maps of actual versus predicted values (BYM Spatial AR(1) model) for the Q1 2017 to Q2 2018 period._](../output/avp_map_series.png)
